---
title: "RNA-Seq Quality Control Report"
author: "Eren Ada, PhD"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
params:
  raw_counts: NULL
  processed_data: NULL
  metadata: NULL
  normality_results: NULL
  pca_results: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 7
)

# Load required packages
library(dplyr)
library(ggplot2)
library(plotly)
library(pheatmap)
library(DT)
library(knitr)
```

# Executive Summary

This report provides a comprehensive quality control analysis of the RNA-Seq dataset. The analysis includes library size distributions, gene detection rates, sample correlations, and normality assessments.

## Dataset Overview

```{r dataset-overview}
# Basic dataset information
n_samples <- ncol(params$raw_counts)
n_genes <- nrow(params$raw_counts)
n_metadata_cols <- ncol(params$metadata)

# Create summary table
overview_df <- data.frame(
  Metric = c("Number of Samples", "Number of Genes", "Number of Metadata Variables"),
  Value = c(n_samples, n_genes, n_metadata_cols)
)

kable(overview_df, caption = "Dataset Overview")
```

# Library Size Analysis

## Distribution of Library Sizes

```{r library-sizes}
# Calculate library sizes
raw_lib_sizes <- colSums(params$raw_counts)
processed_lib_sizes <- colSums(params$processed_data)

# Create plot data
lib_sizes_df <- data.frame(
  Sample = rep(names(raw_lib_sizes), 2),
  LibrarySize = c(raw_lib_sizes, processed_lib_sizes),
  DataType = rep(c("Raw", "Processed"), each = length(raw_lib_sizes))
)

# Create plot
p <- ggplot(lib_sizes_df, aes(x = Sample, y = LibrarySize, fill = DataType)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Library Size Distribution",
       y = "Library Size",
       x = "Sample")

ggplotly(p)
```

## Library Size Statistics

```{r lib-size-stats}
# Calculate statistics
lib_stats <- data.frame(
  Statistic = c("Minimum", "Maximum", "Mean", "Median", "Standard Deviation"),
  Raw = c(min(raw_lib_sizes), max(raw_lib_sizes), mean(raw_lib_sizes),
          median(raw_lib_sizes), sd(raw_lib_sizes)),
  Processed = c(min(processed_lib_sizes), max(processed_lib_sizes),
                mean(processed_lib_sizes), median(processed_lib_sizes),
                sd(processed_lib_sizes))
)

kable(lib_stats, caption = "Library Size Statistics", digits = 2)
```

# Gene Detection Analysis

## Gene Detection Rates

```{r gene-detection}
# Calculate detection rates (genes with counts > 0)
detection_rates <- colMeans(params$processed_data > 0) * 100

# Create plot
detection_df <- data.frame(
  Sample = names(detection_rates),
  DetectionRate = detection_rates
)

p <- ggplot(detection_df, aes(x = Sample, y = DetectionRate)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Gene Detection Rates",
       y = "Detection Rate (%)",
       x = "Sample")

ggplotly(p)
```

# Sample Similarity Analysis

## Principal Component Analysis

```{r pca-plot}
# Extract PCA results
pc_data <- as.data.frame(params$pca_results$pca_object$x)
var_explained <- params$pca_results$var_explained

# Create PCA plot
p <- ggplot(pc_data, aes(x = PC1, y = PC2)) +
  geom_point(size = 3) +
  theme_bw() +
  labs(
    title = "Principal Component Analysis",
    x = sprintf("PC1 (%.1f%%)", var_explained[1]),
    y = sprintf("PC2 (%.1f%%)", var_explained[2])
  )

ggplotly(p)
```

## Sample Correlation Heatmap

```{r correlation-heatmap}
# Calculate correlation matrix
cor_matrix <- cor(params$processed_data)

# Create heatmap
pheatmap(
  cor_matrix,
  main = "Sample Correlation Heatmap",
  color = colorRampPalette(c("blue", "white", "red"))(100)
)
```

# Normality Assessment

## Summary of Normality Tests

```{r normality-summary}
# Extract normality results
sample_stats <- params$normality_results$sample_stats

# Create summary table
normality_df <- do.call(rbind, lapply(names(sample_stats), function(sample) {
  data.frame(
    Sample = sample,
    Skewness = sample_stats[[sample]]$skewness,
    Kurtosis = sample_stats[[sample]]$kurtosis,
    Shapiro_P = sample_stats[[sample]]$shapiro_p
  )
}))

kable(normality_df, caption = "Normality Test Results", digits = 3)
```

## Recommendation

```{r normality-recommendation}
cat(params$normality_results$recommendation)
```

# Quality Control Metrics Summary

```{r qc-summary}
# Create summary of key QC metrics
qc_summary <- data.frame(
  Metric = c(
    "Median Library Size",
    "Mean Detection Rate",
    "Mean Sample Correlation",
    "Samples Passing Normality",
    "PC1 Variance Explained"
  ),
  Value = c(
    median(processed_lib_sizes),
    mean(detection_rates),
    mean(cor_matrix[upper.tri(cor_matrix)]),
    sum(normality_df$Shapiro_P > 0.05, na.rm = TRUE),
    var_explained[1]
  )
)

kable(qc_summary, caption = "Summary of Key QC Metrics", digits = 2)
```

# Session Information

```{r session-info}
sessionInfo()
``` 