---
title: "`r params$report_content$title`"
author: "`r params$report_content$author`"
date: "`r params$report_content$date`"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
params:
  report_content: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(knitr)
library(DT)
library(ggplot2)
```

## Data Processing Summary

### Current Data State
The data is currently in the **`r toupper(params$report_content$data_state)`** state.

### Data Dimensions
- Number of genes: **`r params$report_content$data_summary$dimensions[1]`**
- Number of samples: **`r params$report_content$data_summary$dimensions[2]`**

### Count Statistics
- Total counts: **`r format(params$report_content$data_summary$total_counts, scientific = FALSE, big.mark = ",")`**
- Mean counts per gene: **`r round(params$report_content$data_summary$mean_counts, 2)`**
- Median counts per gene: **`r round(params$report_content$data_summary$median_counts, 2)`**

```{r filtering_params, eval=!is.null(params$report_content$filtering_params)}
if (!is.null(params$report_content$filtering_params)) {
  cat("## Filtering Parameters\n")
  cat("\n### Applied Filters\n")
  cat("- Minimum samples per gene:", params$report_content$filtering_params$min_samples, "\n")
  cat("- Minimum counts per gene:", params$report_content$filtering_params$min_counts, "\n")
  if (params$report_content$filtering_params$group_column != "None") {
    cat("- Grouping column:", params$report_content$filtering_params$group_column, "\n")
  }
}
```

```{r normalization_params, eval=!is.null(params$report_content$normalization_params)}
if (!is.null(params$report_content$normalization_params)) {
  cat("## Normalization\n")
  cat("\n### Method\n")
  cat("Normalization method applied:", params$report_content$normalization_params$method, "\n")
  
  if (!is.null(params$report_content$normalization_params$evaluation_results)) {
    cat("\n### Evaluation Results\n")
    eval_results <- params$report_content$normalization_params$evaluation_results
    # Add evaluation results visualization here if available
  }
}
```

## Session Information
```{r}
sessionInfo()
``` 